Goal Recognition as Planning with Unknown Domain Models

### 2019-02-11 doc/ijcai-19-goalRecognition/goalRecognition.tex
processing Miquels comments for pages 1 and 2 ijcai19 paper on goal recognition

\begin{abstract}
This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em that the observer knows the action model of the observed agents}. The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}. We show that this novel formulation allow us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the action model of the observed agents.  
\end{abstract}

\section{Introduction}
\label{sec:introduction}
{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents pursuing one of those goals. While there exist a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.

{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to estimate the most likely goal of the agents being observed. In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.  In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known beforehand) and it shows that this formulation neatly fits into the {\em plan recognition as planning} approach. The experimental results demonstrate that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without the requirement of having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents

\section{Background}
\label{sec:background}
This section formalizes the {\em planning model} we follow, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.

@@

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\subseteq\mathcal{L}(F)$ is the set of goal conditions over the state variables. A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ being the {\em length} of $\pi$ and $cost(\pi)=\sum_{a\in\pi} cost(a)$ being its {\em cost}. The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$. A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met. A solution plan is {\em optimal} iff its cost is minimal.

@@

\subsection{The observation model}
We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation). Therefore, transiting between two consecutive observed states may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite. In other words, having an input observation does not imply knowing the actual length of the corresponding plan.

@@

\subsection{Goal recognition with classical planning}
{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve a goal $G\in G[\cdot]$.

Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
\begin{align}
argmax_{G\in G[\cdot]} P(\mathcal{O}|G) P(G).
\end{align}

@@

The higher the value of the $cost(\pi^\top)-cost(\pi^\bot)$ difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$. With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:

\begin{align}
P(\mathcal{O}|G) = \frac{1}{1+e^{-\beta(cost(\pi^\top)-cost(\pi^\bot))}}
\end{align}

@@

\section{Planning with unknown domain models}
\label{sec:planning}
This section introduces a novel formulation for classical planning in a setting where no action model is given. This setting has already shown related to the learning of action models for planning~\cite{SternJ17}. In particular it can be seen as an extreme scenario when the action model is learned from a single example that contains only two state observations: the initial state and the goals. A {\em classical planning with unknown domain models} is then a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e. the functions $\rho$ and/or $\theta$ of $a$ are undefined).

A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g. one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state $s_n$ meets the goal condition, $G \subseteq s_n$.

Next we show that the space of possible STRIPS action models can be encoded as a set of propositional variables and a set of constraints over those variables. Then, we show how to exploit this encoding to solve $P=\tup{F,A[\cdot],I,G}$ problems with an off-the-shelf classical planner and the properties of this approach.

\subsection{A propositional encoding for the space of STRIPS action models}
{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema. Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.

For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.

Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of two kinds:
\begin{enumerate}
\item {\bf Syntactic constraints}. \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$. Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$. {\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}. 
\item {\bf Observation constraints}. The observation of the actions and states resulting from the execution of a plan depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
\end{enumerate}

In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$). This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning. In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$. If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$. Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).

@@

In addition, one can introduce {\em domain-specific knowledge} to constrain further the space of possible schemata. For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself. {\it State invariants} are {\em domain-specific knowledge}that can be exploited either as {\em syntactic} constrains to reduce the space of possible action models but also as {\em semantic} constraints to complete partial observations of the states traversed by a plan.

\subsection{A classical planning compilation for planning with unknown domain models}
Now we show how we adapt the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning} to address the task of {\em planning with unknown domain models}, using our propositional encoding of STRIPS action models.

Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
\begin{itemize}
\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models. This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.

\item $A'$ replaces the actions in $A$ with two types of actions.
\begin{enumerate}
\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect into the action model $\xi$ following the syntactic constraints of \strips\ models. 
\begin{itemize}
\item Actions to insert an $e\in{\mathcal I}_{\Psi,\xi}$ {\em precondition} into $\xi$. The precondition is only inserted when neither $pre\_e\_\xi$ nor $eff\_e\_\xi$ exist in $\xi$.

@@

\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$). The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
\end{enumerate}
\end{itemize}

@@

Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation. In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table and the robot hand is empty. The problem goal is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}. The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), steps $07-13$ insert the preconditions and effects of the {\tt\small pickup} action model and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$. Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em putdown} and {\em unstack} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.

@@

The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$. The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$. The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions. Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.

>>>>

\begin{abstract}
This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em the observer knows the action model of the observed agents}. The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}. The experimental results demonstrate that this novel formulation for classical planning allow us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the action model of the observed agents.  
\end{abstract}

\section{Introduction}
\label{sec:introduction}
{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents pursuing one of those goals. While there exist diverse approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2010probabilistic} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.

{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to estimate the most likely goal of the agents being observed~\cite{ramirez2012plan}. In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.  We introduce a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach. The experimental results demonstrate that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents

\section{Background}
\label{sec:background}
This section formalizes the {\em classical planning} model that we follow in this work, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.

@@

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\subseteq\mathcal{L}(F)$ is the set of goal conditions over the state variables. A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ being the {\em length} of $\pi$ and $cost(\pi)=\sum_{a\in\pi} cost(a)$ being its {\em cost}. The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, then $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$. A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met. A solution plan is {\em optimal} iff its cost is minimal.

@@

\subsection{The observation model}
We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation). This means that transiting between two consecutive observed states may require the execution of more than a single action. In other words, having an input observation does not imply knowing the actual length of the corresponding plan.

@@

\subsection{Goal recognition with classical planning}
{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents an observation of agents acting to achieve a goal $G\in G[\cdot]$.

Following the paradigm of the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
\begin{align}
argmax_{G\in G[\cdot]} P(\mathcal{O}|G) P(G).
\end{align}

@@

The higher the value of the $\Delta(\pi^\top,\pi^\bot)$ cost difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$. With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:

\begin{align}
P(\mathcal{O}|G) = \frac{1}{1+e^{-\beta \Delta(\pi^\top,\pi^\bot)}}
\end{align}

@@

\section{Planning with unknown domain models}
\label{sec:planning}
This section introduces a novel formulation for classical planning in a setting where no action model is given. This setting has already shown related to the learning of action models for planning~\cite{SternJ17} and it can be seen as an extreme scenario when the action model is {\em learned} from a single example that contains only two state observations: the initial state and the goals.

A {\em classical planning with unknown domain models} is a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e. the functions $\rho$ and/or $\theta$ of $a$ are undefined). A {\em solution plan} is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g. one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state $s_n$ meets the goal condition $G \subseteq s_n$.

Next we show that a propositional encoding for the space of STRIPS action models allows us to solve $P=\tup{F,A,I,G[\cdot]}$ problems with off-the-shelf classical planners.

\subsection{A propositional encoding for the space of STRIPS action models}
{\em A \strips\ action schema} $\xi$ is defined by: A list of {\em parameters} $pars(\xi)$, and three sets of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema. Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, and is denoted as ${\mathcal I}_{\Psi,\xi}$.

For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the case of the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.

Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints:
\begin{enumerate}
\item {\bf Syntactic constraints}. \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$. Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$. {\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}. 
\item {\bf Observation constraints}. The observation of the actions and states resulting from the execution of a plan depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
\end{enumerate}

In this work we introduce a novel propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ that uses only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$). This encoding exploits the syntactic constraints of \strips\, so it is more compact than previous encodings proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ actions with classical planning. In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$. If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$. To illustrate this, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\em blocksworld} {\tt\small stack(?v1,?v2)} schema and our propositional representation for this schema using only {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).

@@

{\em Domain-specific knowledge} is also helpfull to constrain further the space of possible schemata. For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself. With this regard, {\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.

\subsection{A classical planning compilation for planning with unknown domain models}
Inspired by the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning}, we define here a compilation to address the task of {\em planning with unknown domain models} that uses our compact propositional encoding of STRIPS action models.

Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
\begin{itemize}
\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being {\em programmed} or {\em applied}, and the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents for the propositional encoding of the corresponding space of STRIPS action models.

\item $A'$ replaces the actions in $A$ with two types of actions.
\begin{enumerate}
\item Actions for {\em inserting} a {\em precondition}, {\em positive}/{\em negative} effect into the action model $\xi$ following the syntactic constraints of \strips\ . 
\begin{itemize}
\item Actions to insert an $e\in{\mathcal I}_{\Psi,\xi}$ {\em precondition} into $\xi$. The precondition is only inserted when neither $pre\_e\_\xi$ nor $eff\_e\_\xi$ exist in $\xi$.

@@

\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$). Note that the action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
\end{enumerate}
\end{itemize}

@@

Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation. In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table and the robot hand is empty. The problem goal is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}. The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $05-10$ insert the action model effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} action model (while steps $10-13$ insert its effects) and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$. Another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em unstack} and {\em putdown} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.

@@

The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$. The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$. The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
